---
title: "Functional Frailty AFT with beta(s) — Simulation Report"
author: "Weijia Qian"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 13
    fig_height: 7.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 150)
library(kableExtra)
library(RColorBrewer)
library(tidyverse)

theme_set(
  theme_bw(base_size = 15) +
    theme(legend.position   = "bottom",
          panel.grid.minor  = element_blank(),
          strip.background  = element_rect(fill = "grey92", colour = NA),
          strip.text        = element_text(face = "bold", size = 13),
          plot.subtitle     = element_text(size = 12),
          legend.text       = element_text(size = 13),
          legend.title      = element_text(size = 13))
)

set2 <- brewer.pal(8, "Set2")

# Colour encodes the MODEL throughout: every figure in this report is a
# two-model comparison.
model_cols <- c("Functional frailty AFT" = set2[1],
                "Linear functional Cox (fcoxFr)" = set2[2])

# Set `results_dir` explicitly to pin a run; otherwise the most recent
# outputs/YYYYMMDD folder is used (excluding the interaction study's).
results_dir <- NULL
if (is.null(results_dir)) {
  cand <- list.dirs(here::here("outputs"), recursive = FALSE)
  cand <- cand[grepl("/[0-9]{8}$", cand)]
  stopifnot("No outputs/YYYYMMDD folder found." = length(cand) > 0)
  results_dir <- cand[order(basename(cand), decreasing = TRUE)][1]
}
fig_dir <- here::here("analysis", "figures_noint")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data, cache=TRUE, cache.extra=list(results_dir, length(list.files(results_dir)))}
# Each .RDA holds list(info, coef, s_grid, beta_true, beta).  beta is a list of
# N_iter x nS matrices; only info and coef feed the metric figures.
load_result <- function(path) { e <- new.env(); load(path, envir = e); e$result }

rda_files <- list.files(results_dir, pattern = "\\.RDA$", full.names = TRUE)
raw <- map_dfr(rda_files, function(p) {
  r <- load_result(p)
  if (nrow(r$coef) == 0L) return(NULL)      # scenario where every iteration failed
  cbind(r$info[, setdiff(names(r$info), "seed")], r$coef)
})

n_grid_total <- length(rda_files)
n_scen_have  <- if (nrow(raw)) dplyr::n_distinct(raw$scenario) else 0L

# A runner that loses every iteration still writes a file per scenario, so an
# empty `raw` is the signature of a failed run rather than a missing one.
if (n_scen_have == 0L) {
  stop("No completed iterations in ", results_dir, ": all ", n_grid_total,
       " scenario files contain an empty `coef`.  Every iteration was skipped by ",
       "the runner's tryCatch, which still saves a file, so the run looks ",
       "successful to SLURM.  Check main_simulations.err before re-knitting.")
}

# This report is written for the 81-scenario design (3 families x 3 N_total x
# 3 tau x 3 censoring, J fixed).  Output from the earlier 27-scenario grid, or
# from before the comparator's AFT-scale columns were added, cannot produce
# these figures.  Fail with the specific reason rather than an empty panel.
need <- c("beta_ise", "lf_beta_ise", "tau2_bias", "lf_tau2_bias",
          "c_index_known", "c_index_known_marg", "lf_c_index_known",
          "lf_c_index_known_marg", "ibs_known", "ibs_known_marg",
          "lf_ibs_known", "lf_ibs_known_marg")
missing_cols <- setdiff(need, names(raw))
thin <- c(N_total = dplyr::n_distinct(raw$N_total),
          censor_rate = dplyr::n_distinct(raw$censor_rate),
          tau = dplyr::n_distinct(raw$tau))
if (length(missing_cols) || any(thin < 2)) {
  stop("`", basename(results_dir), "` cannot support this report.\n",
       if (length(missing_cols))
         paste0("  Missing columns: ", paste(missing_cols, collapse = ", "), "\n") else "",
       if (any(thin < 2))
         paste0("  Single-valued design factors: ",
                paste(names(thin)[thin < 2], collapse = ", "),
                " -- the figures facet and pool over these.\n") else "",
       "  Re-run main_simulations.R (81 scenarios) and point results_dir at the new folder.")
}

dat <- raw |>
  mutate(
    family = factor(family, levels = c("lognormal", "cox", "cox_tv"),
                    labels = c("lognormal", "Cox / Weibull PH", "Cox, non-PH")),
    N_f       = factor(N_total, levels = sort(unique(N_total))),
    tau_f     = factor(tau, levels = sort(unique(tau))),
    tau_lab   = factor(paste0("tau = ", tau), levels = paste0("tau = ", sort(unique(tau)))),
    cens_lab  = factor(paste0(100 * censor_rate, "% censoring"),
                       levels = paste0(100 * sort(unique(censor_rate)), "% censoring"))
  )

avail_bt <- sort(unique(dat$beta_type))
```

# Simulation settings

```{r tbl-settings}
grid_levels <- function(v) paste(sort(unique(raw[[v]])), collapse = ", ")

tibble::tribble(
  ~Element, ~Setting, ~Note,
  "Families (DGM)", "lognormal, Cox (Weibull PH), Cox with a time-varying effect",
    "Error shape and, for the last arm, the PH assumption itself",
  "Coefficient function", grid_levels("beta_type"), "B-spline beta(s); no age interaction",
  "Sample size", grid_levels("N_total"), "Cluster sizes ~ Uniform(mean +/- 0.5 mean)",
  "Clusters J", grid_levels("n_cluster"), "Fixed in this design",
  "Frailty SD tau", grid_levels("tau"), "AFT (log-time) scale, common to all families",
  "Censoring rate", grid_levels("censor_rate"), "C ~ Uniform(0, ub), ub solved to hit the target",
  "Residual SD sigma", grid_levels("sigma"),
    "Log-time residual SD target; each DGM takes a different scale parameter",
  "Functional grid nS", grid_levels("nS"), "X(s) via Karhunen-Loeve, Fourier basis, decay k^-0.7",
  "Scenarios", sprintf("%d of %d", n_scen_have, n_grid_total),
    "3 families x 3 N x 3 tau x 3 censoring",
  "Iterations per scenario", as.character(max(raw$iter)), "",
  "Evaluation horizon", sprintf("tau_eval = %.3f - %.3f over cells",
    min(raw$tau_eval), max(raw$tau_eval)),
    "Quantiles of the TRUE event-time distribution, so it is COMMON across censoring levels",
  "Test sets", "N = 1000 marginal (new clusters), N = 1000 conditional (known clusters)",
    "Marginal set fixed within a scenario; conditional set redrawn each iteration",
  "MCMC", "20,000 iterations, 10,000 burn-in, no thinning", "",
  "Priors", "gamma ~ N(0,100); sigma^2 ~ IG(3,2); tau^2 ~ IG(0.01,0.01); lambda ~ Gamma(1,1)",
    "tau^2 prior weakened from IG(3,2), whose prior mean of 1 swamped the frailty likelihood at small tau"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Because the test set is seeded on the *cell* (every design factor except the
censoring rate), the three censoring levels of a cell share byte-identical true
event times and differ only in which of them are observed. The censoring
comparison is therefore **paired**, and the evaluation horizon is common across
it by construction.

```{r tbl-timing}
raw |>
  summarise(gibbs_median = median(time), gibbs_max = max(time),
            comparator_median = median(time_lf),
            total_cpu_hours = sum(time + time_lf) / 3600) |>
  kbl(digits = 1) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

# Models compared

```{r tbl-models}
tibble::tribble(
  ~Feature, ~`Functional frailty AFT`, ~`Linear functional Cox (fcoxFr)`,
  "Scale", "log-time (AFT)", "log-hazard (PH)",
  "Functional effect", "integral X(s) beta(s) ds, penalised spline",
    "integral X(s) beta(s) ds, FPCA basis",
  "Cluster frailty", "u_j ~ N(0, tau^2), shared", "Gaussian random intercept via coxme",
  "Baseline", "Gaussian working error, data augmentation for censoring", "Breslow, unspecified",
  "Inference", "Gibbs sampler, full posterior", "Penalised partial likelihood, point estimates only",
  "Implementation", "gibbs_functional_frailty()", "fpcr() in fcoxFr_supp.R"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

The comparator estimates on the log-hazard scale. Where the DGM fixes a Weibull
shape `rho`, the transform `beta_AFT = -beta_Cox/rho`, `tau2_AFT = tau2_Cox/rho^2`
puts both models on the log-time scale and they share an axis directly. **The
lognormal arm has no PH representation and hence no `rho`**, so the comparator's
beta and tau^2 are structurally untransformable there: it is absent from the
lognormal row of figures 1-4 by construction, not through missing data. The
predictive metrics (figures 5-6) are scale free and defined for every family.

`fpcr()` returns point estimates only, and `coxme` supplies no standard error
for the frailty variance, so no coverage figure has a comparator counterpart.
The comparator is used as built, and coverage is reported in the appendix for
the Bayesian model alone.

# Evaluation metrics

```{r tbl-metrics}
tibble::tribble(
  ~Metric, ~Definition, ~Target,
  "ISE of beta(s)", "Mean squared deviation from the true curve over the s grid, per iteration", "0",
  "Bias of tau^2", "Estimate minus the AFT-scale target. Bayesian: posterior MEDIAN (the tau^2 posterior is right-skewed) minus tau2_implied. Comparator: coxme variance / rho^2 minus the same target", "0",
  "Uno's C-index", "IPCW concordance, truncated at the common horizon tau_eval", "1",
  "IBS", "IPCW integrated Brier score over the common tgrid", "0"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r fig-helpers}
# --- Figures 1-4: two-model boxplot, x = total N, facet family x <one factor> ---
# The other design factor is POOLED, not dropped: every iteration at every level
# of it contributes a point to the box.
box_models <- function(d, yvar, lfvar, ylab, facet_col, hline = NULL) {
  s <- d |>
    select(family, N_f, col = all_of(facet_col),
           a = all_of(yvar), b = all_of(lfvar)) |>
    pivot_longer(c(a, b), names_to = "model", values_to = "value") |>
    mutate(model = factor(model, c("a", "b"), names(model_cols)))
  # The comparator's value is NA on the lognormal arm.  Those rows are KEPT so
  # both fill levels exist in every panel and the dodge allocates the same two
  # slots everywhere -- dropping them instead lets the surviving box stretch to
  # the full slot, so the lognormal boxes render at twice the width of the
  # others.  geom_boxplot drops the NAs itself and simply draws nothing.
  p <- ggplot(s, aes(x = N_f, y = value, fill = model))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.3, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.8, preserve = "single")) +
    facet_grid(family ~ col, scales = "free_y") +
    scale_fill_manual(values = model_cols, name = NULL) +
    guides(fill = guide_legend(nrow = 1)) +
    labs(x = "total sample size N", y = ylab)
}

# --- Figures 5-6: two-model barplot, x = tau, facet family x evaluation --------
# Marginal vs conditional are the SAME held-out subjects in KNOWN clusters,
# scored with the frailty forced to 0 and with u_hat_j respectively.  Using the
# new-cluster test set for "marginal" would change the subjects as well as the
# frailty and confound the two.
# Bar = mean over pooled iterations; whiskers = interquartile range.  The IQR
# shows the SPREAD across the pooled conditions (total N, censoring rate, and
# replicate noise), not the precision of the mean -- with ~900 iterations behind
# each bar a CI of the mean would be a hairline and would say nothing about how
# much the metric actually moves across those conditions.
bar_models <- function(d, marg, cond, lf_marg, lf_cond, ylab, hline = NULL,
                       ybase = NULL) {
  s <- d |>
    select(family, tau_f,
           `Functional frailty AFT.marginal (u = 0)`        = all_of(marg),
           `Functional frailty AFT.conditional (u-hat)`     = all_of(cond),
           `Linear functional Cox (fcoxFr).marginal (u = 0)`    = all_of(lf_marg),
           `Linear functional Cox (fcoxFr).conditional (u-hat)` = all_of(lf_cond)) |>
    pivot_longer(-c(family, tau_f), names_to = c("model", "evaluation"),
                 names_sep = "\\.", values_to = "value") |>
    filter(!is.na(value)) |>
    mutate(model      = factor(model, names(model_cols)),
           evaluation = factor(evaluation,
                               c("marginal (u = 0)", "conditional (u-hat)"))) |>
    group_by(family, tau_f, model, evaluation) |>
    summarise(m  = mean(value),
              lo = quantile(value, 0.25),
              hi = quantile(value, 0.75),
              n  = n(), .groups = "drop")
  p <- ggplot(s, aes(x = tau_f, y = m, fill = model))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_col(position = position_dodge(width = 0.8), width = 0.7, alpha = 0.9) +
    geom_errorbar(aes(ymin = lo, ymax = hi), position = position_dodge(width = 0.8),
                  width = 0.2, linewidth = 0.4) +
    facet_grid(family ~ evaluation,
               scales = if (is.null(ybase)) "free_y" else "fixed") +
    scale_fill_manual(values = model_cols, name = NULL) +
    guides(fill = guide_legend(nrow = 1)) +
    # A C-index bar drawn from 0 puts every value in the top fifth of the panel.
    # 0.5 is the metric's null (a coin-flip ranking), so the floor is the
    # meaningful baseline rather than an arbitrary truncation.  coord_cartesian
    # clips the view without dropping data.
    (if (is.null(ybase)) NULL else coord_cartesian(ylim = c(ybase, NA))) +
    labs(x = expression(tau), y = ylab)
}
```

```{r build-figs}
n_per_box <- nrow(dat) / (dplyr::n_distinct(dat$family) *
                          dplyr::n_distinct(dat$N_f) * 3)
n_per_bar <- nrow(dat) / (dplyr::n_distinct(dat$family) * dplyr::n_distinct(dat$tau_f))

fig_keys <- c("ise_by_censoring", "ise_by_tau", "bias_tau2_by_censoring",
              "bias_tau2_by_tau", "c_index", "ibs")

make_figs <- function(bt) {
  stopifnot(length(fig_keys) == 6L)
  d <- filter(dat, beta_type == bt)
  list(
    "1. ISE of beta(s), by censoring rate" =
      box_models(d, "beta_ise", "lf_beta_ise", expression("ISE of " * beta(s)),
                 "cens_lab", hline = 0) +
        labs(subtitle = paste("Pooled over tau. Both models on the AFT scale:",
                              "the comparator's log-hazard beta_hat is mapped by\n",
                              "beta_AFT = -beta_Cox/rho. Absent from the lognormal row,",
                              "which has no rho.")),
    "2. ISE of beta(s), by tau" =
      box_models(d, "beta_ise", "lf_beta_ise", expression("ISE of " * beta(s)),
                 "tau_lab", hline = 0) +
        labs(subtitle = paste("Pooled over the censoring rate.",
                              "Same quantity as figure 1, sliced the other way.")),
    "3. Bias of tau^2, by censoring rate" =
      box_models(d, "tau2_bias", "lf_tau2_bias", expression("bias of " * tau^2),
                 "cens_lab", hline = 0) +
        labs(subtitle = paste("Pooled over tau. Both against the SAME AFT-scale target",
                              "tau2_implied: the Bayesian posterior median, and\n",
                              "the comparator's coxme variance divided by rho^2.",
                              "Red: zero bias.")),
    "4. Bias of tau^2, by tau" =
      box_models(d, "tau2_bias", "lf_tau2_bias", expression("bias of " * tau^2),
                 "tau_lab", hline = 0) +
        labs(subtitle = paste("Pooled over the censoring rate.",
                              "Same quantity as figure 3, sliced the other way.")),
    "5. Uno's C-index" =
      bar_models(d, "c_index_known_marg", "c_index_known",
                 "lf_c_index_known_marg", "lf_c_index_known",
                 "Uno's IPCW C-index", ybase = 0.5) +
        labs(subtitle = sprintf(paste("Known-cluster test set: the same held-out subjects",
                                      "scored with the frailty forced to 0 and with u_hat_j.\n",
                                      "Pooled over total N and censoring rate (%.0f iterations",
                                      "per bar). Bar: mean; whiskers: interquartile range.",
                                      "\ny axis starts at 0.5, the no-discrimination null."), n_per_bar)),
    "6. Integrated Brier score" =
      bar_models(d, "ibs_known_marg", "ibs_known",
                 "lf_ibs_known_marg", "lf_ibs_known",
                 "IPCW integrated Brier score", hline = 0) +
        labs(subtitle = sprintf(paste("Same test set and pooling as figure 5",
                                      "(%.0f iterations per bar).\n",
                                      "Bar: mean; whiskers: interquartile range.",
                                      "Red: zero (perfect prediction)."), n_per_bar))
  )
}

figs <- setNames(lapply(avail_bt, make_figs), avail_bt)

save_fig <- function(p, name, w = 13, h = 7.5)
  ggsave(file.path(fig_dir, paste0(name, ".png")), p, width = w, height = h, dpi = 300)
for (bt in avail_bt)
  walk(seq_along(figs[[bt]]), ~ save_fig(figs[[bt]][[.x]],
       sprintf("fig%d_%s_%s", .x, fig_keys[.x], bt)))
```

```{r emit, results='asis'}
for (bt in avail_bt) {
  cat("\n\n# ", bt, "\n", sep = "")
  for (k in seq_along(figs[[bt]])) {
    cat("\n\n## ", names(figs[[bt]])[k], " {.unnumbered}\n\n", sep = "")
    print(figs[[bt]][[k]])
  }
}
```

# Summary

```{r tbl-summary}
cols <- c(ISE = "beta_ise", `ISE (fcoxFr)` = "lf_beta_ise",
          `Bias tau2` = "tau2_bias", `Bias tau2 (fcoxFr)` = "lf_tau2_bias",
          `C-index` = "c_index_known", `C-index (fcoxFr)` = "lf_c_index_known",
          IBS = "ibs_known", `IBS (fcoxFr)` = "lf_ibs_known")

dat |>
  group_by(family, censor_rate, tau) |>
  summarise(iterations = n(),
            across(all_of(unname(cols)), ~ mean(.x, na.rm = TRUE)), .groups = "drop") |>
  rename(!!!cols) |>
  kbl(digits = 3, caption = "Means over iterations, pooled across total sample size.") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) |>
  collapse_rows(columns = 1:2, valign = "top")
```
