---
title: "Functional Frailty AFT with beta(s) — Simulation Report"
author: "Weijia Qian"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 13
    fig_height: 7.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 150)
library(kableExtra)
library(RColorBrewer)
library(tidyverse)

theme_set(
  theme_bw(base_size = 15) +
    theme(legend.position   = "bottom",
          panel.grid.minor  = element_blank(),
          strip.background  = element_rect(fill = "grey92", colour = NA),
          strip.text        = element_text(face = "bold", size = 13),
          plot.subtitle     = element_text(size = 12),
          legend.text       = element_text(size = 13),
          legend.title      = element_text(size = 13))
)

set2 <- brewer.pal(8, "Set2")

# Two models are compared here, so colour encodes the model.  Cluster count is
# the x axis and tau the columns; N_total and censoring are single-valued in
# this design and are reported in the settings table instead.
model_cols <- c("Functional frailty AFT" = set2[1],
                "Linear functional Cox (fcoxFr)" = set2[2])

# The predictive figures instead contrast the frailty being off vs on, so they
# need a second pair -- warm against cool, so it separates at a glance.
pred_cols <- c("marginal (u = 0)" = set2[2],
               "conditional (û)"  = set2[3])

results_dir <- here::here("outputs", "20260901")
fig_dir     <- here::here("analysis", "figures_noint")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data, cache=TRUE, cache.extra=length(list.files(results_dir))}
# Each .RDA holds list(info, coef, s_grid, beta_true, beta).  beta is a list of
# N_iter x nS matrices; only info and coef are needed for the metric figures.
load_result <- function(path) { e <- new.env(); load(path, envir = e); e$result }

rda_files <- list.files(results_dir, pattern = "\\.RDA$", full.names = TRUE)
raw <- map_dfr(rda_files, function(p) {
  r <- load_result(p)
  if (nrow(r$coef) == 0L) return(NULL)      # scenario where every iteration failed
  # The comparator reports cor(beta_hat, beta) because its coefficient lives on
  # the log-hazard scale.  Compute the same quantity for the Bayesian fit, per
  # iteration, so the two can share an axis.
  ok  <- which(!apply(r$beta$mean, 1, function(x) all(is.na(x))))
  cb  <- rep(NA_real_, nrow(r$coef))
  cb[seq_along(ok)] <- apply(r$beta$mean[ok, , drop = FALSE], 1,
                             function(b) suppressWarnings(cor(b, r$beta_true)))
  cbind(r$info[, setdiff(names(r$info), "seed")], r$coef,
        beta_shape_cor_bayes = cb)
})

n_grid_total <- length(rda_files)
n_scen_have  <- if (nrow(raw)) dplyr::n_distinct(raw$scenario) else 0L

# A runner that loses every iteration still writes a file per scenario, so an
# empty `raw` is the signature of a failed run rather than a missing one.  Say
# so plainly instead of failing later with an opaque error.
if (n_scen_have == 0L) {
  stop("No completed iterations in ", results_dir, ": all ", n_grid_total,
       " scenario files contain an empty `coef`.  Every iteration was skipped by ",
       "the runner's tryCatch, which still saves a file, so the run looks ",
       "successful to SLURM.  Check main_simulations.err for the repeated ",
       "warning text before re-knitting.")
}

has_comparator <- any(grepl("^lf_", names(raw)))

dat <- raw |>
  mutate(
    family = factor(family, levels = c("lognormal", "cox", "cox_tv"),
                    labels = c("lognormal", "Cox / Weibull PH", "Cox, non-PH")),
    J_f     = factor(n_cluster, levels = sort(unique(n_cluster))),
    tau_lab = factor(paste0("tau = ", tau), levels = paste0("tau = ", sort(unique(tau))))
  )

avail_bt <- sort(unique(dat$beta_type))
```

# Simulation settings

```{r tbl-settings}
grid_levels <- function(v) paste(sort(unique(raw[[v]])), collapse = ", ")

tibble::tribble(
  ~Element, ~Setting, ~Note,
  "Families (DGM)", "lognormal, Cox (Weibull PH), Cox with a time-varying effect",
    "Error shape and, for the last arm, the PH assumption itself",
  "Coefficient function", grid_levels("beta_type"),
    "B-spline beta(s); no age interaction",
  "Clusters J", grid_levels("n_cluster"), "Total N fixed",
  "Frailty SD tau", grid_levels("tau"), "AFT (log-time) scale, common to all families",
  "Sample size", grid_levels("N_total"), "Cluster sizes ~ Uniform(mean +/- 0.5 mean)",
  "Censoring rate", grid_levels("censor_rate"), "C ~ Uniform(0, ub), ub solved to hit the target",
  "Residual SD sigma", grid_levels("sigma"),
    "Log-time residual SD target; each DGM takes a different scale parameter",
  "Functional grid nS", grid_levels("nS"), "X(s) via Karhunen-Loeve, Fourier basis, decay k^-0.7",
  "Scenarios", sprintf("%d of %d", n_scen_have, n_grid_total), "3 families x 3 J x 3 tau",
  "Iterations per scenario", as.character(max(raw$iter)), "",
  "Test sets", "N = 1000 marginal (new clusters), N = 1000 conditional (known clusters)",
    "Marginal set fixed within a scenario; conditional set redrawn each iteration",
  "MCMC", "20,000 iterations, 10,000 burn-in, no thinning", "",
  "Priors", "gamma ~ N(0,100); sigma^2 ~ IG(3,2); tau^2 ~ IG(0.01,0.01); lambda ~ Gamma(1,1)",
    "tau^2 prior weakened from IG(3,2), whose prior mean of 1 swamped the frailty likelihood at small tau"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r tbl-timing}
raw |>
  summarise(gibbs_median = median(time), gibbs_max = max(time),
            comparator_median = if (has_comparator) median(time_lf) else NA_real_,
            total_cpu_hours = sum(time + if (has_comparator) time_lf else 0) / 3600) |>
  kbl(digits = 1, caption = "Wall-clock seconds per fit.") |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

# Models compared

```{r tbl-models}
tibble::tribble(
  ~Feature, ~`Functional frailty AFT`, ~`Linear functional Cox (fcoxFr)`,
  "Scale", "log-time (AFT)", "log-hazard (PH)",
  "Functional effect", "integral X(s) beta(s) ds, penalised spline",
    "integral X(s) beta(s) ds, FPCA basis",
  "Cluster frailty", "u_j ~ N(0, tau^2), shared", "Gaussian random intercept via coxme",
  "Baseline", "Gaussian working error, data augmentation for censoring", "Breslow, unspecified",
  "Inference", "Gibbs sampler, full posterior", "Penalised partial likelihood",
  "Implementation", "gibbs_functional_frailty()", "fpcr() in fcoxFr_supp.R"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Because the comparator works on the log-hazard scale and expands `beta(s)` in an
FPCA basis, its coefficient function is not on the same scale as the AFT target.
It is therefore compared on **shape** (correlation with the truth) and on the
predictive metrics, which are scale-free; ISE and interval coverage are reported
for the Bayesian model only.

# Evaluation metrics

```{r tbl-metrics}
tibble::tribble(
  ~Metric, ~Definition, ~Target,
  "ISE of beta(s)", "Mean squared deviation from the true curve over the s grid; the mean over iterations is the MISE", "0",
  "Pointwise coverage", "Fraction of s-grid points whose true value falls in the 95% pointwise CrI", "0.95",
  "Simultaneous coverage", "Whether the ENTIRE true curve falls inside the 95% max-statistic band (one 0/1 per iteration)", "0.95",
  "beta shape correlation", "cor(beta_hat(s), beta(s)) -- the comparator's only scale-free beta summary", "1",
  "Bias of tau^2", "Posterior MEDIAN minus the AFT-scale target (the tau^2 posterior is right-skewed)", "0",
  "Coverage of tau^2", "Whether the target falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "Bias of sigma^2", "Posterior mean minus the log-time residual variance", "0",
  "Coverage of sigma^2", "Whether sigma^2_true falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "C-index / IBS, new clusters", "Uno's IPCW C-index and IPCW integrated Brier score, frailty = 0", "1 / 0",
  "C-index / IBS, known clusters", "The same held-out subjects with the frailty off (u = 0) and on (u_hat)", "1 / 0"
) |>
  kbl(caption = paste("Estimation metrics are scored on the AFT (log-time) scale.",
                      "For the Cox arms the target is the implied AFT quantity, not",
                      "the DGM parameter.")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Figures

Rows are the data-generating family and columns τ; x is the number of clusters.
Total sample size and censoring rate are single-valued in this design, so nothing
is pooled away. Where both models supply a metric, colour separates them.

```{r fig-helpers}
box_panel <- function(d, yvar, ylab, hline = NULL) {
  p <- ggplot(d, aes(x = J_f, y = .data[[yvar]]))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(fill = model_cols[[1]], outlier.size = 0.3, linewidth = 0.3,
                 alpha = 0.9, width = 0.6) +
    facet_grid(family ~ tau_lab) +
    labs(x = "number of clusters J", y = ylab)
}

# Two models on the same axes, where the comparator supplies the metric too.
box_panel2 <- function(d, yvar, lfvar, ylab, hline = NULL) {
  s <- d |>
    select(family, tau_lab, J_f, a = all_of(yvar), b = all_of(lfvar)) |>
    pivot_longer(c(a, b), names_to = "model", values_to = "value") |>
    mutate(model = factor(model, c("a", "b"), names(model_cols)))
  p <- ggplot(s, aes(x = J_f, y = value, fill = model))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.3, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.8)) +
    facet_grid(family ~ tau_lab) +
    scale_fill_manual(values = model_cols, name = NULL) +
    guides(fill = guide_legend(nrow = 1)) +
    labs(x = "number of clusters J", y = ylab)
}

# Coverage indicators are 0/1 per iteration, so the informative summary is the
# rate over iterations with a Clopper-Pearson interval, not a boxplot.
rate_panel <- function(d, cvar, ylab) {
  s <- d |>
    group_by(family, tau_lab, J_f) |>
    summarise(k = sum(.data[[cvar]]), n = n(), .groups = "drop") |>
    mutate(rate = k / n,
           lo = if_else(k == 0, 0, qbeta(0.025, k, n - k + 1)),
           hi = if_else(k == n, 1, qbeta(0.975, k + 1, n - k)))
  ggplot(s, aes(x = J_f, y = rate)) +
    geom_hline(yintercept = 0.95, colour = "red", linetype = "dotted", linewidth = 0.6) +
    geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.18, linewidth = 0.5,
                  colour = model_cols[[1]]) +
    geom_point(size = 2.4, colour = model_cols[[1]]) +
    facet_grid(family ~ tau_lab) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(x = "number of clusters J", y = ylab)
}

# Known-cluster test set scored twice from the same fit and the same held-out
# subjects: frailty off, then frailty set to u_hat_j.  Only the frailty differs,
# so the gap between the colours is what the frailty contributes.
pred_panel <- function(d, marg, cond, ylab, hline = NULL) {
  s <- d |>
    select(family, tau_lab, J_f, marginal = all_of(marg), conditional = all_of(cond)) |>
    pivot_longer(c(marginal, conditional), names_to = "evaluation", values_to = "value") |>
    mutate(evaluation = factor(recode(evaluation,
                                      marginal    = "marginal (u = 0)",
                                      conditional = "conditional (û)"),
                               levels = names(pred_cols)))
  p <- ggplot(s, aes(x = J_f, y = value, fill = evaluation))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.3, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.8)) +
    facet_grid(family ~ tau_lab) +
    scale_fill_manual(values = pred_cols, name = NULL) +
    labs(x = "number of clusters J", y = ylab)
}
```

```{r load-beta, cache=TRUE, dependson="load-data"}
# beta(s) averaged over the completed iterations of each scenario.
beta_dat <- map_dfr(rda_files, function(p) {
  r <- load_result(p)
  if (nrow(r$coef) == 0L) return(NULL)
  i <- r$info[1, ]
  ok <- !apply(r$beta$mean, 1, function(x) all(is.na(x)))
  tibble(family = i$family, beta_type = i$beta_type, tau = i$tau,
         n_cluster = i$n_cluster, s = r$s_grid, beta_true = r$beta_true,
         est  = colMeans(r$beta$mean[ok, , drop = FALSE]),
         q025 = colMeans(r$beta$q025[ok, , drop = FALSE]),
         q975 = colMeans(r$beta$q975[ok, , drop = FALSE]))
})
```

```{r fig-beta-curve}
beta_curve <- function(bt) {
  beta_dat |>
    filter(beta_type == bt) |>
    group_by(family, tau, s) |>                 # average over cluster count
    summarise(across(c(beta_true, est, q025, q975), mean), .groups = "drop") |>
    mutate(family = factor(family, c("lognormal", "cox", "cox_tv"),
                           c("lognormal", "Cox / Weibull PH", "Cox, non-PH")),
           tau_lab = factor(paste0("tau = ", tau))) |>
    ggplot(aes(s)) +
    geom_ribbon(aes(ymin = q025, ymax = q975), fill = model_cols[[1]], alpha = 0.25) +
    geom_line(aes(y = beta_true), linetype = "dotted", linewidth = 0.9, colour = "grey20") +
    geom_line(aes(y = est), colour = model_cols[[1]], linewidth = 1) +
    facet_grid(family ~ tau_lab, scales = "free_y") +
    labs(x = "s", y = expression(hat(beta)(s)),
         subtitle = paste("Dotted grey = truth; line = posterior mean averaged over",
                          "iterations and cluster counts;\nribbon = average pointwise",
                          "95% credible band."))
}
```

```{r build-figs, include=FALSE}
fig_keys <- c("mise_beta", "cover_pointwise_beta", "cover_simultaneous_beta",
              "beta_shape_cor", "bias_tau2", "cover_tau2", "bias_sigma2",
              "cover_sigma2", "c_index", "ibs", "frailty_gain", "beta_curve")

make_figs <- function(bt) {
  # fig_keys names the saved files; a mismatch would silently write fig12_NA.png
  stopifnot(length(fig_keys) == 12L)
  d <- dat |> filter(beta_type == bt)
  list(
    "1. ISE of beta(s)" =
      box_panel(d, "beta_ise", expression("ISE of " * beta(s))) +
        labs(subtitle = "Bayesian model only. Mean over iterations = MISE."),
    "2. Pointwise coverage" =
      box_panel(d, "beta_cover_pw",
                expression("Pointwise 95% CrI coverage of " * beta(s)), hline = 0.95) +
        labs(subtitle = "Fraction of s-grid points covered, per iteration. Red: nominal 0.95."),
    "3. Simultaneous coverage" =
      rate_panel(d, "beta_cover_sim",
                 expression("Simultaneous 95% coverage of " * beta(s))) +
        labs(subtitle = "Whole-curve coverage rate. Red: nominal 0.95."),
    "4. Shape recovery of beta(s)" =
      (if (has_comparator)
         box_panel2(d, "beta_shape_cor_bayes", "lf_beta_cor",
                    expression("cor(" * hat(beta)(s) * ", " * beta(s) * ")"), hline = 1)
       else box_panel(d, "beta_shape_cor_bayes",
                    expression("cor(" * hat(beta)(s) * ", " * beta(s) * ")"), hline = 1)) +
        labs(subtitle = paste("Scale-free, so the log-hazard comparator is comparable here.",
                              "Red: perfect shape.")),
    "5. Bias of tau^2" =
      box_panel(d, "tau2_bias", expression("Bias of " * tau^2), hline = 0) +
        labs(subtitle = "Posterior median minus the AFT-scale target. Red: zero bias."),
    "6. Coverage of tau^2" =
      rate_panel(d, "tau2_cover", expression("95% CrI coverage of " * tau^2)) +
        labs(subtitle = "Coverage rate. Red: nominal 0.95."),
    "7. Bias of sigma^2" =
      box_panel(d, "sigma2_bias", expression("Bias of " * sigma^2), hline = 0) +
        labs(subtitle = "Posterior mean minus the log-time residual variance. Red: zero bias."),
    "8. Coverage of sigma^2" =
      rate_panel(d, "sigma2_cover", expression("95% CrI coverage of " * sigma^2)) +
        labs(subtitle = "Coverage rate. Red: nominal 0.95."),
    "9. C-index" =
      (if (has_comparator)
         box_panel2(d, "c_index", "lf_c_index", "Uno's IPCW C-index")
       else box_panel(d, "c_index", "Uno's IPCW C-index")) +
        labs(subtitle = "Marginal test set: new subjects in new clusters, frailty = 0."),
    "10. Integrated Brier score" =
      (if (has_comparator)
         box_panel2(d, "ibs", "lf_ibs", "IPCW integrated Brier score", hline = 0)
       else box_panel(d, "ibs", "IPCW integrated Brier score", hline = 0)) +
        labs(subtitle = "Marginal test set. Red: zero (perfect prediction)."),
    "11. Frailty contribution" =
      pred_panel(d, "c_index_known_marg", "c_index_known", "Uno's IPCW C-index") +
        labs(subtitle = paste("Known-cluster test set, Bayesian model:",
                              "frailty 0 vs û_j.")),
    "12. Estimated beta(s)" = beta_curve(bt)
  )
}

figs <- setNames(lapply(avail_bt, make_figs), avail_bt)

save_fig <- function(p, name, w = 13, h = 7.5)
  ggsave(file.path(fig_dir, paste0(name, ".png")), p, width = w, height = h, dpi = 300)
for (bt in avail_bt)
  walk(seq_along(figs[[bt]]), ~ save_fig(figs[[bt]][[.x]],
       sprintf("fig%d_%s_%s", .x, fig_keys[.x], bt)))
```

```{r emit, results='asis'}
emit_figures <- function(fs) {
  for (k in seq_along(fs)) {
    cat("\n\n## ", names(fs)[k], " {.unnumbered}\n\n", sep = "")
    print(fs[[k]])
  }
}
for (bt in avail_bt) {
  cat("\n\n# ", bt, "\n", sep = "")
  emit_figures(figs[[bt]])
}
```

# Summary

```{r tbl-summary}
sub <- dat
cols <- c(ISE = "beta_ise", `Cover pw` = "beta_cover_pw", `Cover sim` = "beta_cover_sim",
          `Bias tau2` = "tau2_bias", `Cover tau2` = "tau2_cover",
          `Bias sigma2` = "sigma2_bias", `Cover sigma2` = "sigma2_cover",
          `C-index` = "c_index", IBS = "ibs")
if (has_comparator) cols <- c(cols, `C-index (fcoxFr)` = "lf_c_index",
                              `IBS (fcoxFr)` = "lf_ibs")

sub |>
  group_by(family, tau, n_cluster) |>
  summarise(iterations = n(), across(all_of(unname(cols)), mean), .groups = "drop") |>
  rename(!!!cols) |>
  kbl(digits = 3, caption = "Means over iterations.") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) |>
  collapse_rows(columns = 1:2, valign = "top")
```
