---
title: "Functional Frailty AFT with a Bivariate Surface — Simulation Report"
author: "Weijia Qian"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 11
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 150)
library(kableExtra)
library(RColorBrewer)
library(tidyverse)

theme_set(
  theme_bw(base_size = 15) +
    theme(legend.position   = "bottom",
          panel.grid.minor  = element_blank(),
          strip.background  = element_rect(fill = "grey92", colour = NA),
          strip.text        = element_text(face = "bold", size = 13),
          plot.subtitle     = element_text(size = 12),
          legend.text       = element_text(size = 13),
          legend.title      = element_text(size = 13))
)

set2 <- brewer.pal(8, "Set2")

# This batch fits a single model (the no-interaction comparator is commented out
# in the runner), so colour is free to encode the evaluation instead: the
# predictive figures show the same fit scored with the frailty off and on.
one_col   <- set2[1]
# Cluster count is the only factor pooled: it explains <=2% of the
# between-scenario variance in every metric.  Family, N, tau and censoring all
# stay visible -- censoring as the panel row, tau as the fill.
tau_cols <- setNames(brewer.pal(3, "Set2"), c("0.2", "0.5", "1"))
# Warm vs cool rather than two shades of one hue: the frailty contrast is the
# point of these two figures, so the pair needs to separate at a glance.
pred_cols <- c("marginal (u = 0)" = set2[2],   # orange
               "conditional (û)"  = set2[3])   # blue-violet

results_dir <- here::here("outputs", "20260901_interaction")
fig_dir     <- here::here("analysis", "figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data, cache=TRUE, cache.extra=length(list.files(results_dir))}
# Each .RDA holds list(info, coef, grid, s_eval, surface).  Only info and coef
# are needed here; the surface matrices are read separately below.
load_result <- function(path) { e <- new.env(); load(path, envir = e); e$result }

rda_files <- list.files(results_dir, pattern = "\\.RDA$", full.names = TRUE)
raw <- map_dfr(rda_files, function(p) {
  r <- load_result(p)
  cbind(r$info[, setdiff(names(r$info), "seed")], r$coef)
})

# The no-interaction comparator is commented out in the runner for this batch,
# so there are no ni_* columns and every figure below is single-model.
n_grid_total <- 243   # 3 families x 3 N x 3 J x 1 surface x 3 tau x 3 censoring

dat <- raw |>
  mutate(
    family = factor(family, levels = c("lognormal", "loglogistic", "cox"),
                    labels = c("lognormal", "loglogistic", "Cox / Weibull PH")),
    censor_lab = factor(paste0(100 * censor_rate, "% censoring"),
                        levels = paste0(100 * sort(unique(censor_rate)), "% censoring")),
    N_f   = factor(N_total, levels = sort(unique(N_total))),
    tau_f = factor(tau)
  )
```

```{r pick-subset}
# Figures pool over n_cluster and censor_rate: every point is an average across
# whichever (J, censoring) cells exist for that family / N / tau.
sub <- dat
avail_bt <- sort(unique(sub$beta_type))

# How many scenarios go into each point.  In a complete run this is 9 everywhere
# (3 J x 3 censoring); here it is not, so it is drawn on the panels.
cell_n <- sub |>
  distinct(family, censor_lab, N_f, scenario) |>
  count(family, censor_lab, N_f, name = "n_scen")

# Annotating every panel with the same number is clutter; only label the panels
# when the pooling is uneven and the reader needs to discount some of them.
show_counts <- n_distinct(cell_n$n_scen) > 1
```

```{r load-surfaces, cache=TRUE, dependson="load-data"}
# Fitted surfaces for the estimated-surface figure.  Only the scenarios the
# figures actually use are read, and only the posterior mean averaged over
# replicates is kept -- the full matrices stay on disk.
scen_fig <- dat |> distinct(scenario) |> pull(scenario)

surf_dat <- map_dfr(scen_fig, function(sc) {
  r <- load_result(file.path(results_dir, paste0(sc, ".RDA")))
  i <- r$info[1, ]
  tibble(family = i$family, beta_type = i$beta_type, tau = i$tau,
         n_cluster = i$n_cluster, censor_rate = i$censor_rate,
         s = r$grid$s, age = r$grid$age, beta_true = r$grid$beta_true,
         est = colMeans(r$surface$mean))
})
```

# Simulation settings

```{r tbl-settings}
grid_levels <- function(v) paste(sort(unique(raw[[v]])), collapse = ", ")

tibble::tribble(
  ~Element, ~Setting, ~Note,
  "Families (DGM)", "lognormal, loglogistic, Cox (Weibull PH)",
    "Error shape only; variance and signal matched across arms",
  "Coefficient surfaces", grid_levels("beta_type"),
    "additive s+a; bilinear 2sa; ridge exp(-6(s-a)^2)",
  "Clusters J", grid_levels("n_cluster"), "Total N fixed at 4000",
  "Frailty SD τ", grid_levels("tau"), "AFT (log-time) scale, common to all families",
  "Censoring rate", grid_levels("censor_rate"), "C ~ Uniform(0, ub), ub solved to hit the target",
  "Residual SD σ", grid_levels("sigma"),
    "Log-time residual SD target; each DGM takes a different scale parameter",
  "Sample size", grid_levels("N_total"), "Cluster sizes ~ Uniform(mean±0.5·mean)",
  "Functional grid nS", grid_levels("nS"), "X(s) via Karhunen-Loève, Fourier basis, decay k^-0.7",
  "Scenarios", sprintf("%d of %d", n_distinct(raw$scenario), n_grid_total),
    "3 families x 3 N x 3 J x 1 surface x 3 τ x 3 censoring",
  "Iterations per scenario", as.character(max(raw$iter)), "All scenarios completed",
  "Test sets", "N = 1000 marginal (new clusters), N = 1000 conditional (known clusters)",
    "Marginal set fixed within a scenario; conditional set redrawn each iteration",
  "MCMC", "20,000 iterations, 10,000 burn-in, no thinning", "Identical for both models",
  "Priors", "γ ~ N(0,100); σ² ~ IG(3,2); τ² ~ IG(0.01,0.01); λ ~ Gamma(1,1)",
    "τ² prior weakened from IG(3,2), which had prior mean 1 and swamped the frailty likelihood at small τ"
) |>
 kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r tbl-timing, include=FALSE}
raw |>
  summarise(median = median(time), max = max(time),
            total_cpu_hours = sum(time) / 3600) |>
  kbl(digits = 1, caption = "Gibbs sampler wall-clock seconds per fit.") |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

# Model fitted

```{r tbl-models}
tibble::tribble(
  ~Feature, ~Specification,
  "Functional effect", "∫ X(s) β(s, age) ds",
  "Basis", "Tensor product te(s, age), K_s = 15, K_age = 10, cubic regression",
  "Smoothing parameters", "λ₁ (s), λ₂ (age), both Gamma(1,1)",
  "Cluster frailty", "u_j ~ N(0, τ²), shared",
  "Error", "Gaussian working error, σ² ~ IG(3,2)",
  "Frailty variance prior", "τ² ~ IG(0.01, 0.01)",
  "Censoring", "Data augmentation, truncated normal",
  "Implementation", "gibbs_frailty_interaction()"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Evaluation metrics

## Scoring the Cox arm on the AFT scale

For the Cox DGM, inverting the Weibull baseline
$h_0(t)=\lambda_0\rho t^{\rho-1}$ with $T=\{E/(\lambda_0 e^{\eta})\}^{1/\rho}$,
$E\sim\mathrm{Exp}(1)$, gives

$$\log T \;=\; -\frac{\log\lambda_0}{\rho} \;-\; \frac{\eta_{\text{Cox}}}{\rho}
\;-\; \frac{u_j}{\rho} \;+\; \frac{\log E}{\rho},$$

so every effect maps to log time by division by $-\rho$, where $\rho$ is the Weibull shape parameter 
($\rho = 1$ gives a constant hazard (exponential), $\rho > 1$ a rising one, $\rho < 1$ a falling one).

The log-time error is $(\log E)/\rho \sim \mathrm{Gumbel\_min}(0,1)/\rho$, so 
$\sigma^2_{AFT} = \pi^2 / (6\rho^2)$.

The DGM is specified at $\rho\beta$, $\rho\gamma$ and $\rho\tau$, so that after the
transform the Cox arm sits at exactly the same signal and noise as the AFT arms.

```{r tbl-cox-transform}
cx <- raw |> filter(family == "cox") |> slice(1)
ln <- raw |> filter(family == "lognormal", tau == cx$tau) |> slice(1)

tibble::tribble(
  ~Quantity, ~`Cox DGM (log-hazard)`, ~Transform, ~`AFT target (log-time)`,
  "β(s, a)", "ρ · β(s,a)", "−β_Cox / ρ", "−β(s,a)",
  "γ₁ (slope)", sprintf("%.4f", cx$gamma_true_1), "−γ_Cox / ρ",
    sprintf("%.4f", cx$gamma_implied_1), 
  "γ₀ (intercept)", sprintf("%.4f", cx$gamma_true_0),
    "−(log λ₀ + γ₀ + γ_EM) / ρ", 
    sprintf("%.4f", cx$gamma_implied_0),
  "τ²", sprintf("%.4f", cx$tau2_true), "τ²_Cox / ρ²",
    sprintf("%.4f", cx$tau2_implied), 
  "σ²", "—", "π² / (6ρ²)",
    sprintf("%.4f", cx$sigma2_true), 
) |>
  kbl(caption = sprintf(
    paste("Cox arm at ρ = %.4f, τ = %.1f. The AFT target column is what bias and coverage",
          "are scored against; it matches the AFT arms by construction, except the intercept."), cx$rho, cx$tau)) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Metric definitions

```{r tbl-metrics}
tibble::tribble(
  ~Metric, ~Definition, ~Target,
  "ISE of β(s,a)", "Mean squared deviation from the true surface over a 30x30 (s, age) grid; the mean over iterations is the MISE", "0",
  "Pointwise coverage", "Fraction of grid points whose true value falls in the 95% pointwise CrI", "0.95",
  "Simultaneous coverage", "Whether the ENTIRE true surface falls inside the 95% max-statistic band (one 0/1 per iteration)", "0.95",
  "Bias of τ²", "Posterior MEDIAN minus the AFT-scale target τ²_implied (the τ² posterior is right-skewed, so the mean sits above the truth)", "0",
  "Coverage of τ²", "Whether τ²_implied falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "Bias of σ²", "Posterior mean minus the log-time residual variance", "0",
  "Coverage of σ²", "Whether σ²_true falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "C-index / IBS — new clusters", "`c_index`, `ibs`: new subjects in NEW clusters, frailty = 0 (population-level prediction)", "1 / 0",
  "C-index / IBS — known clusters, u = 0", "`c_index_known_marg`, `ibs_known_marg`: held-out subjects in KNOWN clusters, frailty still set to 0", "1 / 0",
  "C-index / IBS — known clusters, û", "`c_index_known`, `ibs_known`: the same held-out subjects, scored with the estimated frailty û_j", "1 / 0",
) |>
  kbl(caption = "All estimation metrics are scored on the AFT (log-time) scale.") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Figures

Results are **pooled over cluster count only** — 3 scenarios and
`r 3 * max(dat$iter)` fits behind each box. Rows are the censoring rate and
columns the data-generating family; x is the total sample size and colour is τ,
so all four factors that matter stay visible. Cluster count is pooled because it
explains ≤ 2% of the between-scenario variance in every metric.

The two predictive figures spend colour on the frailty contrast (frailty off vs
on), so there x is τ and N is pooled instead.

```{r fig-helpers}
# Every panel builder shares this layout: x = total sample size, columns = tau,
# rows = family, pooled over n_cluster and censor_rate.  The count of pooled
# scenarios is printed at the top of each panel.
add_counts <- function(p, d, yvar) {
  if (!show_counts) return(p)
  lab <- cell_n |> semi_join(distinct(d, family, censor_lab, N_f),
                             by = c("family", "censor_lab", "N_f"))
  yr  <- range(d[[yvar]], na.rm = TRUE)
  p +
    geom_text(data = lab, aes(x = N_f, y = yr[2] + 0.10 * diff(yr), label = n_scen),
              inherit.aes = FALSE, size = 2.7, colour = "grey35") +
    expand_limits(y = yr[2] + 0.15 * diff(yr))
}

box_panel <- function(d, yvar, ylab, hline = NULL) {
  p <- ggplot(d, aes(x = N_f, y = .data[[yvar]], fill = tau_f))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p <- p +
    geom_boxplot(outlier.size = 0.3, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.8)) +
    facet_grid(family ~ censor_lab) +
    scale_fill_manual(values = tau_cols, name = expression(tau)) +
    labs(x = "total sample size N", y = ylab)
  add_counts(p, d, yvar)
}

# Coverage indicators are 0/1 per iteration, so the informative summary is the
# rate over iterations with a Clopper-Pearson interval, not a boxplot.
rate_panel <- function(d, cvar, ylab) {
  s <- d |>
    group_by(family, tau_f, N_f, censor_lab) |>
    summarise(k = sum(.data[[cvar]]), n = n(), .groups = "drop") |>
    mutate(rate = k / n,
           lo = if_else(k == 0, 0, qbeta(0.025, k, n - k + 1)),
           hi = if_else(k == n, 1, qbeta(0.975, k + 1, n - k)))
  p <- ggplot(s, aes(x = N_f, y = rate, colour = tau_f)) +
    geom_hline(yintercept = 0.95, colour = "red", linetype = "dotted", linewidth = 0.6) +
    geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, linewidth = 0.5,
                  position = position_dodge(width = 0.6)) +
    geom_point(size = 2.1, position = position_dodge(width = 0.6)) +
    facet_grid(family ~ censor_lab) +
    scale_colour_manual(values = tau_cols, name = expression(tau)) +
    scale_y_continuous(limits = c(0, 1.08)) +
    labs(x = "total sample size N", y = ylab)
  if (!show_counts) return(p)
  lab <- cell_n |> semi_join(distinct(s, family, censor_lab, N_f),
                             by = c("family", "censor_lab", "N_f"))
  p + geom_text(data = lab, aes(x = N_f, y = 1.06, label = n_scen),
                inherit.aes = FALSE, size = 2.7, colour = "grey35")
}

# Predictive metrics on the KNOWN-cluster test set, scored twice from the same
# fit and the same held-out subjects: once with the frailty set to 0 and once
# with the estimated frailty u_hat_j.  Only the frailty differs, so the gap
# between the two shades is exactly what the frailty contributes.
pred_panel <- function(d, marg, cond, ylab, hline = NULL) {
  s <- d |>
    select(family, tau_f, censor_lab, marginal = all_of(marg),
           conditional = all_of(cond)) |>
    pivot_longer(c(marginal, conditional), names_to = "evaluation",
                 values_to = "value") |>
    mutate(evaluation = factor(recode(evaluation,
                                      marginal    = "marginal (u = 0)",
                                      conditional = "conditional (û)"),
                               levels = names(pred_cols)))
  p <- ggplot(s, aes(x = tau_f, y = value, fill = evaluation))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.4, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.8)) +
    facet_grid(family ~ censor_lab) +
    scale_fill_manual(values = pred_cols, name = NULL) +
    labs(x = expression(tau), y = ylab)
}

# Estimated surface as slices in s at three ages, one panel per family.  A heat
# map would need a shared fill scale across families, and the Cox arm's
# AFT-scale surface is NEGATIVE by construction (beta_AFT = -beta_Cox/rho), so
# slices with a free y per family read better.
slice_ages <- function() {
  a <- sort(unique(surf_dat$age))
  vapply(c(40, 60, 80), function(x) a[which.min(abs(a - x))], numeric(1))
}
age_cols <- setNames(brewer.pal(9, "BuGn")[c(5, 7, 9)],
                     sprintf("age %g", round(slice_ages())))

surface_slice <- function(bt) {
  keep <- slice_ages()
  surf_dat |>
    filter(beta_type == bt, age %in% keep) |>
    group_by(family, censor_rate, s, age) |>  # average over n_cluster, tau and N
    summarise(across(c(beta_true, est), mean), .groups = "drop") |>
    mutate(family = factor(family, c("lognormal", "loglogistic", "cox"),
                           c("lognormal", "loglogistic", "Cox / Weibull PH")),
           censor_lab = factor(paste0(100 * censor_rate, "% censoring")),
           series = factor(sprintf("age %g", round(age)), levels = names(age_cols))) |>
    ggplot(aes(s, colour = series)) +
    geom_line(aes(y = beta_true, group = series), linetype = "dotted", linewidth = 0.9) +
    geom_line(aes(y = est, group = series), linewidth = 1) +
    facet_grid(family ~ censor_lab, scales = "free_y") +
    scale_colour_manual(values = age_cols, name = NULL) +
    labs(x = "s", y = expression(hat(beta)(s, a)),
         subtitle = paste0(
           "Dotted = truth at each age; solid = posterior mean averaged over ",
           "replicates, cluster counts,\ntau and sample sizes.  ",
           "The Cox AFT surface is negative by construction."))
}

fig_keys   <- c("mise_beta", "cover_pointwise_beta", "cover_simultaneous_beta",
                "bias_tau2", "cover_tau2", "bias_sigma2", "cover_sigma2",
                "c_index", "ibs", "surface_fit")
fig_widths  <- c(rep(13, 7), 12, 12, 11)
fig_heights <- c(rep(7, 9), 8)

save_fig <- function(p, name, w = 11, h = 7)
  ggsave(file.path(fig_dir, paste0(name, ".png")), p, width = w, height = h, dpi = 300)

surface_figures <- function(bt, save = TRUE) {
  d <- sub |> filter(beta_type == bt)
  fs <- list(
    "1. ISE of the surface" =
      box_panel(d, "beta_ise", expression("ISE of " * beta(s, a))) +
        labs(subtitle = "Mean over iterations = MISE."),
    "2. Pointwise coverage" =
      box_panel(d, "beta_cover_pw",
                expression("Pointwise 95% CrI coverage of " * beta(s, a)), hline = 0.95) +
        labs(subtitle = "Fraction of the 900 grid points covered, per iteration. Red: nominal 0.95."),
    "3. Simultaneous coverage" =
      rate_panel(d, "beta_cover_sim",
                 expression("Simultaneous 95% coverage of " * beta(s, a))) +
        labs(subtitle = "Whole-surface coverage rate. Red: nominal 0.95."),
    "4. Bias of tau^2" =
      box_panel(d, "tau2_bias", expression("Bias of " * tau^2), hline = 0) +
        labs(subtitle = "Posterior median minus the AFT-scale target. Red: zero bias."),
    "5. Coverage of tau^2" =
      rate_panel(d, "tau2_cover", expression("95% CrI coverage of " * tau^2)) +
        labs(subtitle = "Coverage rate. Red: nominal 0.95."),
    "6. Bias of sigma^2" =
      box_panel(d, "sigma2_bias", expression("Bias of " * sigma^2), hline = 0) +
        labs(subtitle = "Posterior mean minus the log-time residual variance. Red: zero bias."),
    "7. Coverage of sigma^2" =
      rate_panel(d, "sigma2_cover", expression("95% CrI coverage of " * sigma^2)) +
        labs(subtitle = "Coverage rate. Red: nominal 0.95."),
    "8. C-index" =
      pred_panel(d, "c_index_known_marg", "c_index_known", "Uno's IPCW C-index") +
        labs(subtitle = paste("Known-cluster test set: frailty 0 vs û_j.",
                              "Pooled over total sample size and cluster count.")),
    "9. Integrated Brier score" =
      pred_panel(d, "ibs_known_marg", "ibs_known",
                 "IPCW integrated Brier score", hline = 0) +
        labs(subtitle = paste("Known-cluster test set: frailty 0 vs û_j.",
                              "Pooled over total sample size and cluster count.",
                              "Red: zero.")),
    "10. Estimated surface" = surface_slice(bt)
  )
  if (save)
    walk(seq_along(fs), ~ save_fig(fs[[.x]],
         sprintf("fig%d_%s_%s", .x, fig_keys[.x], bt),
         w = fig_widths[.x], h = fig_heights[.x]))
  fs
}

emit_figures <- function(fs, idx) {
  for (k in idx) {
    cat("\n\n## ", names(fs)[k], " {.unnumbered}\n\n", sep = "")
    print(fs[[k]])
  }
}
```

```{r build-figs, include=FALSE}
figs <- setNames(lapply(avail_bt, surface_figures), avail_bt)
```

`r if (length(avail_bt) > 1) "Sections follow for each coefficient surface present in the batch." else ""`

```{r emit, results='asis', fig.width=13, fig.height=7.5}
for (bt in avail_bt) {
  cat("\n\n# ", bt, " surface\n", sep = "")
  emit_figures(figs[[bt]], seq_along(figs[[bt]]))
}
```

# Summary of the plotted subset

```{r tbl-summary}
sub |>
  group_by(family, tau, N_total) |>
  summarise(
    scenarios      = n_distinct(scenario),
    iterations     = n(),
    ISE            = mean(beta_ise),
    `Cover pw`     = mean(beta_cover_pw),
    `Cover sim`    = mean(beta_cover_sim),
    `Bias tau2`    = mean(tau2_bias),
    `Cover tau2`   = mean(tau2_cover),
    `Bias sigma2`  = mean(sigma2_bias),
    `Cover sigma2` = mean(sigma2_cover),
    `C known u=0`  = mean(c_index_known_marg),
    `C known û`    = mean(c_index_known),
    `IBS known û`  = mean(ibs_known),
    .groups = "drop"
  ) |>
  kbl(digits = 3, caption = paste("Means over iterations, pooled over cluster",
                                  "count and censoring rate.")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) |>
  collapse_rows(columns = 1:2, valign = "top")
```
