---
title: "Functional Frailty AFT with a Bivariate Surface — Simulation Report"
author: "Weijia Qian"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 11
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 150)
library(kableExtra)
library(RColorBrewer)
library(tidyverse)

theme_set(
  theme_bw(base_size = 12) +
    theme(legend.position   = "bottom",
          panel.grid.minor  = element_blank(),
          strip.background  = element_rect(fill = "grey92", colour = NA),
          strip.text        = element_text(face = "bold", size = 10))
)

set2 <- brewer.pal(8, "Set2")
model_cols <- c("Interaction β(s,a)"  = set2[1],
                "No interaction β(s)" = set2[2])

# For the predictive figures each model appears twice (marginal / conditional),
# so the second evaluation reuses the model's hue at 60% luminance: same colour
# family, still separable.
shade <- function(hex, f) {
  m <- col2rgb(hex) * f
  rgb(m[1], m[2], m[3], maxColorValue = 255)
}
pred_cols <- c(
  "Interaction β(s,a) — marginal (u = 0)"     = model_cols[[1]],
  "Interaction β(s,a) — conditional (û)"      = shade(model_cols[[1]], 0.60),
  "No interaction β(s) — marginal (u = 0)"    = model_cols[[2]],
  "No interaction β(s) — conditional (û)"     = shade(model_cols[[2]], 0.60)
)

results_dir <- here::here("outputs", "20260826")
fig_dir     <- here::here("analysis", "figures")
dir.create(fig_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data, cache=TRUE, cache.extra=tools::md5sum(list.files(results_dir, full.names=TRUE)[1])}
# Each .RDA holds list(info, coef, grid, s_eval, surface, ni_surface).  Only
# info and coef are needed here; the surface matrices are left on disk.
load_result <- function(path) { e <- new.env(); load(path, envir = e); e$result }

rda_files <- list.files(results_dir, pattern = "\\.RDA$", full.names = TRUE)
raw <- map_dfr(rda_files, function(p) {
  r <- load_result(p)
  cbind(r$info[, setdiff(names(r$info), "seed")], r$coef)
})

# Metrics that exist for BOTH models (the comparator carries an ni_ prefix).
paired_metrics <- c(
  "beta_ise", "beta_ise_rel", "beta_bias_mean", "beta_bias_max",
  "beta_cover_pw", "beta_cover_sim", "beta_cma_qd",
  "tau2_est", "tau2_bias", "tau2_se", "tau2_cover",
  "sigma2_est", "sigma2_bias", "sigma2_se", "sigma2_cover",
  "c_index", "ibs", "c_index_known", "ibs_known",
  "c_index_known_marg", "ibs_known_marg"
)
design_cols <- c("scenario", "iter", "family", "N_total", "n_cluster", "nS",
                 "beta_type", "tau", "tau_dgm", "sigma", "sigma_dgm", "rho",
                 "censor_rate", "event_rate", "time", "time_ni")

# Long by model, so every figure is one geom with a colour aesthetic.
dat <- bind_rows(
  raw |> select(all_of(design_cols), all_of(paired_metrics)) |>
    mutate(model = "Interaction β(s,a)"),
  raw |> select(all_of(design_cols), all_of(paste0("ni_", paired_metrics))) |>
    rename_with(~ sub("^ni_", "", .x), starts_with("ni_")) |>
    mutate(model = "No interaction β(s)")
) |>
  mutate(
    model  = factor(model, levels = names(model_cols)),
    family = factor(family, levels = c("lognormal", "loglogistic", "cox"),
                    labels = c("lognormal\n(correctly specified)",
                               "loglogistic\n(heavy tails)",
                               "Cox / Weibull PH\n(skewed)")),
    censor_lab = factor(paste0(100 * censor_rate, "% censoring"),
                        levels = paste0(100 * c(0.25, 0.5, 0.75), "% censoring")),
    tau_f = factor(tau)
  )

# Figures below use this subset.
sub <- dat |> filter(n_cluster == 50, beta_type == "additive")
```

```{r load-surfaces, cache=TRUE, dependson="load-data"}
# The fitted surfaces themselves, for the estimated-surface figure.  Only the
# n_cluster = 50 scenarios are read (81 of 243 files), and only the posterior
# mean averaged over replicates is kept -- the full matrices stay on disk.
scen50 <- raw |> filter(n_cluster == 50) |> distinct(scenario) |> pull(scenario)

surf_dat <- map_dfr(scen50, function(sc) {
  r <- load_result(file.path(results_dir, paste0(sc, ".RDA")))
  i <- r$info[1, ]
  tibble(family = i$family, beta_type = i$beta_type,
         tau = i$tau, censor_rate = i$censor_rate,
         s = r$grid$s, age = r$grid$age, beta_true = r$grid$beta_true,
         est_int = colMeans(r$surface$mean),
         est_ni  = colMeans(r$ni_surface$mean)[match(r$grid$s, r$s_eval)])
})
```

# Simulation settings

```{r tbl-settings}
grid_levels <- function(v) paste(sort(unique(raw[[v]])), collapse = ", ")

tibble::tribble(
  ~Element, ~Setting, ~Note,
  "Families (DGM)", "lognormal, loglogistic, Cox (Weibull PH)",
    "Error shape only; variance and signal matched across arms",
  "Coefficient surfaces", grid_levels("beta_type"),
    "additive s+a; bilinear 2sa; ridge exp(-6(s-a)^2)",
  "Clusters J", grid_levels("n_cluster"), "Total N fixed at 4000",
  "Frailty SD τ", grid_levels("tau"), "AFT (log-time) scale, common to all families",
  "Censoring rate", grid_levels("censor_rate"), "C ~ Uniform(0, ub), ub solved to hit the target",
  "Residual SD σ", grid_levels("sigma"),
    "Log-time residual SD target; each DGM takes a different scale parameter",
  "Sample size", grid_levels("N_total"), "Cluster sizes ~ Uniform(mean±0.5·mean)",
  "Functional grid nS", grid_levels("nS"), "X(s) via Karhunen-Loève, Fourier basis, decay k^-0.7",
  "Scenarios", as.character(n_distinct(raw$scenario)), "3 families x 3 J x 3 surfaces x 3 τ x 3 censoring",
  "Iterations per scenario", as.character(nrow(raw) / n_distinct(raw$scenario)), "All scenarios completed",
  "Test sets", "N = 1000 marginal (new clusters), N = 1000 conditional (known clusters)",
    "Marginal set fixed within a scenario; conditional set redrawn each iteration",
  "MCMC", "20,000 iterations, 10,000 burn-in, no thinning", "Identical for both models",
  "Priors", "γ ~ N(0,100); σ², τ² ~ IG(3,2); λ ~ Gamma(1,1)", "Identical for both models"
) |>
 kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r tbl-timing, include=FALSE}
raw |>
  summarise(across(c(time, time_ni), list(median = median, max = max))) |>
  mutate(total_cpu_hours = sum(raw$time + raw$time_ni) / 3600) |>
  kbl(digits = 1, caption = "Sampler wall-clock seconds per fit (time = interaction, time_ni = comparator).") |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

# Models compared

```{r tbl-models}
tibble::tribble(
  ~Feature, ~`Interaction model`, ~`No-interaction model`,
  "Functional effect", "∫ X(s) β(s, age) ds", "∫ X(s) β(s) ds",
  "Basis", "Tensor product te(s, age), K_s = 15, K_age = 10, cubic regression",
    "Marginal spline in s only, K = 15, cubic regression (the tensor's s-margin)",
  "Smoothing parameters", "λ₁ (s), λ₂ (age), both Gamma(1,1)", "λ, Gamma(1,1)",
  # "Cluster frailty", "u_j ~ N(0, τ²), shared", "u_j ~ N(0, τ²), shared",
  # "Error", "Gaussian working error, σ² ~ IG(3,2)", "Gaussian working error, σ² ~ IG(3,2)",
  # "Censoring", "Data augmentation, truncated normal", "Data augmentation, truncated normal",
  # "Fitted to", "The same training data", "The same training data",
  # "Scored on", "The same grid, test sets and targets", "β̂(s) broadcast along age",
  "Implementation", "gibbs_frailty_interaction()", "gibbs_functional_frailty()"
) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Evaluation metrics

## Scoring the Cox arm on the AFT scale

For the Cox DGM, inverting the Weibull baseline
$h_0(t)=\lambda_0\rho t^{\rho-1}$ with $T=\{E/(\lambda_0 e^{\eta})\}^{1/\rho}$,
$E\sim\mathrm{Exp}(1)$, gives

$$\log T \;=\; -\frac{\log\lambda_0}{\rho} \;-\; \frac{\eta_{\text{Cox}}}{\rho}
\;-\; \frac{u_j}{\rho} \;+\; \frac{\log E}{\rho},$$

so every effect maps to log time by division by $-\rho$, where $\rho$ is the Weibull shape parameter 
($\rho = 1$ gives a constant hazard (exponential), $\rho > 1$ a rising one, $\rho < 1$ a falling one).

The log-time error is $(\log E)/\rho \sim \mathrm{Gumbel\_min}(0,1)/\rho$, so 
$\sigma^2_{AFT} = \pi^2 / (6\rho^2)$.

The DGM is specified at $\rho\beta$, $\rho\gamma$ and $\rho\tau$, so that after the
transform the Cox arm sits at exactly the same signal and noise as the AFT arms.

```{r tbl-cox-transform}
cx <- raw |> filter(family == "cox", beta_type == "additive", n_cluster == 50) |> slice(1)
ln <- raw |> filter(family == "lognormal", beta_type == "additive", n_cluster == 50,
                    tau == cx$tau) |> slice(1)

tibble::tribble(
  ~Quantity, ~`Cox DGM (log-hazard)`, ~Transform, ~`AFT target (log-time)`,
  "β(s, a)", "ρ · β(s,a)", "−β_Cox / ρ", "−β(s,a)",
  "γ₁ (slope)", sprintf("%.4f", cx$gamma_true_1), "−γ_Cox / ρ",
    sprintf("%.4f", cx$gamma_implied_1), 
  "γ₀ (intercept)", sprintf("%.4f", cx$gamma_true_0),
    "−(log λ₀ + γ₀ + γ_EM) / ρ", 
    sprintf("%.4f", cx$gamma_implied_0),
  "τ²", sprintf("%.4f", cx$tau2_true), "τ²_Cox / ρ²",
    sprintf("%.4f", cx$tau2_implied), 
  "σ²", "—", "π² / (6ρ²)",
    sprintf("%.4f", cx$sigma2_true), 
) |>
  kbl(caption = sprintf(
    paste("Cox arm at ρ = %.4f, τ = %.1f. The AFT target column is what bias and coverage",
          "are scored against; it matches the AFT arms by construction, except the intercept."), cx$rho, cx$tau)) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Metric definitions

```{r tbl-metrics}
tibble::tribble(
  ~Metric, ~Definition, ~Target,
  "ISE of β(s,a)", "Mean squared deviation from the true surface over a 30x30 (s, age) grid; the mean over iterations is the MISE", "0",
  "Pointwise coverage", "Fraction of grid points whose true value falls in the 95% pointwise CrI", "0.95",
  "Simultaneous coverage", "Whether the ENTIRE true surface falls inside the 95% max-statistic band (one 0/1 per iteration)", "0.95",
  "Bias of τ²", "Posterior mean minus the AFT-scale target τ²_implied", "0",
  "Coverage of τ²", "Whether τ²_implied falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "Bias of σ²", "Posterior mean minus the log-time residual variance", "0",
  "Coverage of σ²", "Whether σ²_true falls in the 95% CrI (one 0/1 per iteration)", "0.95",
  "C-index / IBS — new clusters", "`c_index`, `ibs`: new subjects in NEW clusters, frailty = 0 (population-level prediction)", "1 / 0",
  "C-index / IBS — known clusters, u = 0", "`c_index_known_marg`, `ibs_known_marg`: held-out subjects in KNOWN clusters, frailty still set to 0", "1 / 0",
  "C-index / IBS — known clusters, û", "`c_index_known`, `ibs_known`: the same held-out subjects, scored with the estimated frailty û_j", "1 / 0",
) |>
  kbl(caption = "All estimation metrics are scored on the AFT (log-time) scale.") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Additive surface {.tabset}

All figures in this section use **J = 50 clusters** and the **additive** surface
(`r n_distinct(sub$scenario)` scenarios, `r nrow(sub)/2/n_distinct(sub$scenario)` iterations each).
Columns are the data-generating family, rows the censoring rate; the y-axis is
fixed across panels.

```{r fig-helpers}
box_panel <- function(d, yvar, ylab, hline = NULL, hline_lab = NULL) {
  p <- ggplot(d, aes(x = tau_f, y = .data[[yvar]], fill = model))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.5, linewidth = 0.35, alpha = 0.85,
                 position = position_dodge(width = 0.8)) +
    facet_grid(censor_lab ~ family) +
    scale_fill_manual(values = model_cols, name = NULL) +
    labs(x = expression(tau), y = ylab)
}

# Coverage indicators are 0/1 per iteration, so the informative summary is the
# rate over iterations with a Clopper-Pearson interval.
rate_panel <- function(d, cvar, ylab) {
  s <- d |>
    group_by(family, censor_lab, tau_f, model) |>
    summarise(k = sum(.data[[cvar]]), n = n(), .groups = "drop") |>
    mutate(rate = k / n,
           lo = if_else(k == 0, 0, qbeta(0.025, k, n - k + 1)),
           hi = if_else(k == n, 1, qbeta(0.975, k + 1, n - k)))
  ggplot(s, aes(x = tau_f, y = rate, colour = model)) +
    geom_hline(yintercept = 0.95, colour = "red", linetype = "dotted", linewidth = 0.6) +
    geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.18,
                  position = position_dodge(width = 0.55), linewidth = 0.5) +
    geom_point(size = 2.2, position = position_dodge(width = 0.55)) +
    facet_grid(censor_lab ~ family) +
    scale_colour_manual(values = model_cols, name = NULL) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(x = expression(tau), y = ylab)
}

# Predictive metrics on the KNOWN-cluster test set, scored twice from the same
# fit and the same held-out subjects: once with the frailty set to 0 (marginal)
# and once with the estimated frailty u_hat_j (conditional).  Because only the
# frailty differs, the gap between the two shades is exactly what the frailty
# contributes -- no confounding from the test sets differing.
pred_panel <- function(d, marg, cond, ylab, hline = NULL) {
  s <- d |>
    select(family, censor_lab, tau_f, model,
           marginal = all_of(marg), conditional = all_of(cond)) |>
    pivot_longer(c(marginal, conditional),
                 names_to = "evaluation", values_to = "value") |>
    mutate(evaluation = recode(evaluation,
                               marginal    = "marginal (u = 0)",
                               conditional = "conditional (û)"),
           series = factor(paste0(model, " — ", evaluation),
                           levels = names(pred_cols)))
  p <- ggplot(s, aes(x = tau_f, y = value, fill = series))
  if (!is.null(hline))
    p <- p + geom_hline(yintercept = hline, colour = "red",
                        linetype = "dotted", linewidth = 0.6)
  p +
    geom_boxplot(outlier.size = 0.4, linewidth = 0.3, alpha = 0.9,
                 position = position_dodge(width = 0.85)) +
    facet_grid(censor_lab ~ family) +
    scale_fill_manual(values = pred_cols, name = NULL) +
    guides(fill = guide_legend(nrow = 2)) +
    labs(x = expression(tau), y = ylab)
}

save_fig <- function(p, name, w = 11, h = 7)
  ggsave(file.path(fig_dir, paste0(name, ".png")), p, width = w, height = h, dpi = 300)

# Estimated surface, shown as slices in s at three ages, all in one panel per
# family. 
slice_ages <- function() {
  a <- sort(unique(surf_dat$age))
  vapply(c(40, 60, 80), function(x) a[which.min(abs(a - x))], numeric(1))
}

age_cols <- setNames(brewer.pal(9, "BuGn")[c(5, 7, 9)],   # lightest = Set2 teal
                     sprintf("age %g", round(slice_ages())))
slice_cols <- c(age_cols, "no-interaction fit" = unname(model_cols[2]))

surface_slice <- function(bt) {
  keep   <- slice_ages()
  fam_lb <- c("lognormal", "loglogistic", "Cox / Weibull PH")

  base <- surf_dat |>
    filter(beta_type == bt, age %in% keep) |>
    group_by(family, s, age) |>                      # average over tau and censoring
    summarise(across(c(beta_true, est_int, est_ni), mean), .groups = "drop") |>
    mutate(family = factor(family, c("lognormal", "loglogistic", "cox"), fam_lb),
           series = factor(sprintf("age %g", round(age)), levels = names(slice_cols)))

  # the comparator is identical at every age, so keep a single curve
  ni <- base |>
    filter(age == keep[1]) |>
    mutate(series = factor("no-interaction fit", levels = names(slice_cols)))

  ggplot(mapping = aes(s, colour = series)) +
    geom_line(data = base, aes(y = beta_true, group = series),
              linetype = "dotted", linewidth = 0.9) +
    geom_line(data = base, aes(y = est_int, group = series), linewidth = 1) +
    geom_line(data = ni, aes(y = est_ni, group = series), linewidth = 1) +
    facet_wrap(~ family, nrow = 1, scales = "free_y") +
    scale_colour_manual(values = slice_cols, name = NULL) +
    labs(x = "s", y = expression(hat(beta)(s, a)),
         subtitle = paste0(
           "Dotted = truth at each age; solid = posterior mean averaged over 100 replicates ",
           "and all tau /\ncensoring levels (J = 50).  The no-interaction fit is one curve, ",
           "the same at every age.\nThe Cox AFT surface is negative by construction."))
}

# One builder for the nine figures, so the three surfaces stay in sync.
# Returns a named list; names double as the section headings.
fig_keys   <- c("mise_beta", "cover_pointwise_beta", "cover_simultaneous_beta",
                "bias_tau2", "cover_tau2", "bias_sigma2", "cover_sigma2",
                "c_index", "ibs", "surface_fit")
fig_widths <- c(rep(11, 7), 13, 13, 12)
fig_heights<- c(rep(7, 9), 4.6)

surface_figures <- function(bt, save = TRUE) {
  d   <- dat |> filter(n_cluster == 50, beta_type == bt)
  lab <- sprintf("J = 50, %s surface", bt)
  fs <- list(
    "1. ISE of the surface" =
      box_panel(d, "beta_ise", expression("ISE of " * beta(s, a))) +
        labs(subtitle = paste0(lab, ". Mean over iterations = MISE.")),
    "2. Pointwise coverage" =
      box_panel(d, "beta_cover_pw",
                expression("Pointwise 95% CrI coverage of " * beta(s, a)), hline = 0.95) +
        labs(subtitle = "Fraction of the 900 grid points covered, per iteration. Red: nominal 0.95."),
    "3. Simultaneous coverage" =
      rate_panel(d, "beta_cover_sim",
                 expression("Simultaneous 95% coverage of " * beta(s, a))) +
        labs(subtitle = "Whole-surface coverage rate over 100 iterations. Red: nominal 0.95."),
    "4. Bias of $\\tau^2$" =
      box_panel(d, "tau2_bias", expression("Bias of " * tau^2), hline = 0) +
        labs(subtitle = "Posterior mean minus the AFT-scale target. Red: zero bias."),
    "5. Coverage of $\\tau^2$" =
      rate_panel(d, "tau2_cover", expression("95% CrI coverage of " * tau^2)) +
        labs(subtitle = "Coverage rate over 100 iterations. Red: nominal 0.95."),
    "6. Bias of $\\sigma^2$" =
      box_panel(d, "sigma2_bias", expression("Bias of " * sigma^2), hline = 0) +
        labs(subtitle = "Posterior mean minus the log-time residual variance. Red: zero bias."),
    "7. Coverage of $\\sigma^2$" =
      rate_panel(d, "sigma2_cover", expression("95% CrI coverage of " * sigma^2)) +
        labs(subtitle = "Coverage rate over 100 iterations. Red: nominal 0.95."),
    "8. C-index" =
      pred_panel(d, "c_index_known_marg", "c_index_known", "Uno's IPCW C-index") +
        labs(subtitle = "Known-cluster test set: frailty set to 0 vs frailty set to û_j."),
    "9. Integrated Brier score" =
      pred_panel(d, "ibs_known_marg", "ibs_known",
                 "IPCW integrated Brier score", hline = 0) +
        labs(subtitle = "Known-cluster test set: frailty 0 vs û_j. Red: zero (perfect prediction)."),
    "10. Estimated surface" = surface_slice(bt)
  )
  if (save)
    walk(seq_along(fs), ~ save_fig(fs[[.x]],
         sprintf("fig%d_%s_%s", .x, fig_keys[.x], bt),
         w = fig_widths[.x], h = fig_heights[.x]))
  fs
}

# Emit a run of figures under their own headings (used by the bilinear and
# ridge sections, which do not need per-figure prose).
emit_figures <- function(fs, idx, level = "##") {
  walk(idx, function(i) {
    cat("\n\n", level, " ", names(fs)[i], "\n\n", sep = "")
    print(fs[[i]])
    cat("\n\n")
  })
}

summary_table <- function(bt) {
  dat |>
    filter(n_cluster == 50, beta_type == bt) |>
    group_by(censor_lab, family, model) |>
    summarise(
      ISE            = mean(beta_ise),
      `Cover pw`     = mean(beta_cover_pw),
      `Cover sim`    = mean(beta_cover_sim),
      `Bias tau2`    = mean(tau2_bias),
      `Cover tau2`   = mean(tau2_cover),
      `Bias sigma2`  = mean(sigma2_bias),
      `Cover sigma2` = mean(sigma2_cover),
      `C new clust`  = mean(c_index),
      `C known u=0`  = mean(c_index_known_marg),
      `C known û`    = mean(c_index_known),
      `IBS known u=0`= mean(ibs_known_marg),
      `IBS known û`  = mean(ibs_known),
      .groups = "drop"
    ) |>
    kbl(digits = 3,
        caption = sprintf("Means over tau and iterations, J = 50, %s surface.", bt)) |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) |>
    collapse_rows(columns = 1:2, valign = "top")
}

figs_additive <- surface_figures("additive")
figs_bilinear <- surface_figures("bilinear")
figs_ridge    <- surface_figures("ridge")
```

## 1. ISE of the surface

```{r fig1-mise}
figs_additive[[1]]
```

## 2. Pointwise coverage

```{r fig2-cover-pw}
figs_additive[[2]]
```

## 3. Simultaneous coverage

Coverage rate over the 100 iterations, with a Clopper–Pearson 95% interval.

```{r fig3-cover-sim}
figs_additive[[3]]
```

## 4. Bias of $\tau^2$

```{r fig4-tau2-bias}
figs_additive[[4]]
```

## 5. Coverage of $\tau^2$

```{r fig5-tau2-cover}
figs_additive[[5]]
```

## 6. Bias of $\sigma^2$

```{r fig6-sigma2-bias}
figs_additive[[6]]
```

## 7. Coverage of $\sigma^2$

```{r fig7-sigma2-cover}
figs_additive[[7]]
```

## 8. C-index

The **same** held-out subjects in **known** clusters are used.
The only difference is whether the estimated frailty û_j is applied.

```{r fig8-cindex, fig.width=13}
figs_additive[[8]]
```

## 9. Integrated Brier score

```{r fig9-ibs, fig.width=13}
figs_additive[[9]]
```

## 10. Estimated surface

```{r fig10-surface-additive, fig.width=12, fig.height=4.6}
figs_additive[[10]]
```

## Summary

```{r tbl-summary}
sub |>
  group_by(censor_lab, family, model) |>
  summarise(
    ISE            = mean(beta_ise),
    `Cover pw`     = mean(beta_cover_pw),
    `Cover sim`    = mean(beta_cover_sim),
    `Bias tau2`    = mean(tau2_bias),
    `Cover tau2`   = mean(tau2_cover),
    `Bias sigma2`  = mean(sigma2_bias),
    `Cover sigma2` = mean(sigma2_cover),
    `C new clust`  = mean(c_index),
    `C known u=0`  = mean(c_index_known_marg),
    `C known û`    = mean(c_index_known),
    `IBS known u=0`= mean(ibs_known_marg),
    `IBS known û`  = mean(ibs_known),
    .groups = "drop"
  ) |>
  kbl(digits = 3, caption = "Means over tau and iterations, J = 50, additive surface.") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) |>
  collapse_rows(columns = 1:2, valign = "top")
```

# Bilinear surface {.tabset}

$\beta(s,a) = 2sa$ — a pure multiplicative interaction, zero along $s=0$ or $a=0$. Same layout and subset as above (J = 50).

```{r figs-bilinear-1, results='asis', fig.width=11, fig.height=7}
emit_figures(figs_bilinear, 1:7)
```

```{r figs-bilinear-2, results='asis', fig.width=13, fig.height=7}
emit_figures(figs_bilinear, 8:9)
```

```{r figs-bilinear-3, results='asis', fig.width=12, fig.height=4.6}
emit_figures(figs_bilinear, 10)
```

## Summary

```{r tbl-summary-bilinear}
summary_table("bilinear")
```

# Ridge surface {.tabset}

$\beta(s,a) = \exp\{-6(s-a)^2\}$ — a diagonal ridge that is genuinely
non-separable. Same layout and subset as above (J = 50).

```{r figs-ridge-1, results='asis', fig.width=11, fig.height=7}
emit_figures(figs_ridge, 1:7)
```

```{r figs-ridge-2, results='asis', fig.width=13, fig.height=7}
emit_figures(figs_ridge, 8:9)
```

```{r figs-ridge-3, results='asis', fig.width=12, fig.height=4.6}
emit_figures(figs_ridge, 10)
```

## Summary

```{r tbl-summary-ridge}
summary_table("ridge")
```
